### 2025-09-25 paper/sec/authoring-workflow.tex
Contribs.

\subsection{\InterpretationAgent}
\label{sec:interpretation-agent}

\cdnote{I integrated the prompt description in the description of the InterpretationAgentv (unsure if this is the best plece) -- the prompt description seemed to basic (it's commented if you want to revert)}

The InterpretationAgent is guided by a structured system prompt that
frames code generation as a precise replacement task. The model
receives the imported datasets, helper modules, and the current Fluid
representation of the paragraph, in which a text fragment is marked
with [REPLACE …]. Its task is to substitute this placeholder with a
Fluid expression that evaluates exactly to the target string,
reconstructing quantitative or comparative claims as dataset queries
where appropriate. To ensure integration with the workflow, the output
must consist solely of a syntactically valid Fluid expression, with no
additional commentary. This focused design constrains the model while
providing enough context to generate expressions that are both
data-grounded and immediately executable; the full prompt is given in~\appref{system-prompt:interpretation-agent}.

>>>>

\paragraph{\InterpretationAgent prompt design.}

The \InterpretationAgent is guided by a structured system prompt that frames code generation as a precise
replacement task. The model receives the imported datasets, helper modules, and the current Fluid
representation of the paragraph, in which a text fragment is marked with the tag \kw{[REPLACE …]}. Its task is
to substitute this placeholder with a Fluid expression that evaluates exactly to the target string,
reconstructing quantitative or comparative claims as data queries. To ensure integration with the workflow,
the output must consist solely of a syntactically valid Fluid expression, with no additional commentary. The
full prompt is given in~\appref{system-prompt:interpretation-agent}.

### 2025-09-25 paper/sec/related-work.tex
Add some notes

As we saw earlier \todo{ref}, scientific texts routinely make use of comparatives like ``faster'' while
leaving one of the argument slots implicit, with the context determining the omitted referent. LLMs
demonstrate considerable competence in resolving these and other more syntactic forms of anaphora such as
pronouns~\citep{zhu25}, but the resolved referent itself -- concretely, what was being referred to -- remains
implicit. Interpretable NLP is a recent research direction which aims to support comprehension (and
production) of text in a more explicit and transparent way~\citep{yulan23}. By generating code that formalises
the interpretation of a comparative like ``faster'', our approach also makes these implicit references
explicit; a richer version of our approach would allow the user to explore the linguistic interpretation as
well. \todo{WIP}

>>>>

As \figref{scigen-example-website} illustrates, scientific texts routinely make use of comparatives like
``faster'' while leaving one of the argument slots implicit, with the context determining the omitted
referent. LLMs demonstrate considerable competence in resolving these and other more syntactic forms of
anaphora such as pronouns~\citep{zhu25}, but the resolved referent itself -- concretely, what was being
referred to -- remains implicit. Interpretable NLP is a recent research direction which aims to support
comprehension (and production) of text in a more explicit and transparent way~\citep{yulan23}. By generating
code that formalises the interpretation of a comparative like ``faster'', our approach also makes these
implicit references explicit; combining our system with interpretable NLP would allow the user to explore the
linguistic interpretation as well.
